\textbf{\sc Towards Deliberative Control of Autonomous Robots}

### 2012-02-14 arch.tex
updates

\item \emph{Partitioning} the global planning problem into multiple
  decision loops, called {\em reactors}. Each reactor has its own
  scope both functionally (one for example focusing on the conversion
  of high level waypoints into low-level commands) and temporally as
  each reactor declares its planning look-ahead and expected latency
  before being able to produce plan to the agent depending on where it
  is located in a reactor dependency graph such as shown in
  Fig. \ref{fig:agent}.

@@

@@

\item \texttt{step} that accepts new goals requested (if any) on the
  {\em internal} timeline of a reactor and will execute one step of
  deliberation. In the context of \eu for example pone step of
  deliberation correspod to resolve one sibngle flaw while ensuring
  that the resulting plan is not proven inconsistent. This implies
  that it also includes the backtracking of the search in such case. 
  The returned value by this step function is a boolean which
  indicates whether the reactor needs some extra step to deliberate or
  if this last step resulted on a complete plan for the reactor
  look-ahead. On the later, the part of the plan that concern this
  reactor {\em external} timelines is provided in order to be
  dispatched as goals to whichever reactor that is reponsible for
  these. \comment{By extending I may have saud things I developped
    below ... will have  to check}
\end{itemize}

@@

For the sake of simplicity we introduce first how a reactor can track
and identify its state in the context where it does not have any
compelling need to deliberate. Assume here that we have a reactor that
have no future goal. Even in this context, it is a core requirement in
T-REX (see \cite{py10}) to have each reactor to produce at every
single tick their {\em internal} state or at least, for the ones not owning
{\em internal} timelines, make sure that {\em external} observations
are taken into account for every tick. This requirement is enforced to
ensure that all the reactors of an agent share a consistent view of
the world as time advance.

Taking into account {\em external} state updates is manage d by
introducing these observation in the plan structure as factsand ensure
that they remain compatible sewith the current plan mainainted by the
reactor. The more delicate part of synchronization is to decide of the 
{\em internla} state values for the current tick. In the Shopping
example for example we have the following code snippet whihch helps
identify when we can \texttt{Buy} a product:
\begin{verbatim}
Agent::Buy {
  eq(duration, 10);
  // [...]
  contained_by(condition object.location.At currLocation);
  // [...]
}
\end{verbatim}
If \texttt{object.location} is {\em external} and the \texttt{Agent}
is {\em internal}. I may encounter the situation where I am at the
location where I can buy one (possibly among others) product. Already
there's here  a decision problem related to the state value of the
\texttt{Agent} timeline where I can either {\em Buy} a specific
product now, wait a little longer before doing so or just simply not
buy any product at this location.

In the \eu based reactor we introduce this decision reuqirement 
as a new type of flaw for the \eu framework. This flaw enforces at the
reactor to identify fully its internal state for the current tick. By
using this new flaw we can describe the synchronization process as
this general sequence :

@@

We introduce this requirement as a new type of flaw (Section
\ref{sec:europa:arch}) for the \eu framework. This flaw forces the
reactor to identify its internal state for the current tick. By using
this flaw one can describe the synchronization process as:

@@

While this model is perfectly standard and acceptable in the overall
case. It can become problematic when tied to execution. Specifically it
puts a strong tie between the {\em fact} that we are \texttt{Going} to a
location and the {\em expected} future outcome of ending at this
location. This ties is even stronger due to the constraint at line 5.
Knowing this aspect assume that the \texttt{Agent.location} is {\em
  external} to the reactor which receive the observation
\texttt{Going(Home,SuperMarket)} which synchronization resolved by
producing the observation \texttt{Go(Home, SuperMarket)} in the Agent
timeline. From this point this means that the only possible outcome on
\texttt{Agent.location} is to be \texttt{At(SuperMarket)} in the
directly foreseeable future. In the real world the fact that we are
\texttt{Going} to a certain location does not necessarly imply that we
will succeed on ending up there. An extreme case could be that due to
some exogenous event, the road to the \texttt{SuperMarket} is closed
which could result on an alternate observation \texttt{At(7th st. \& Foo
  Ave.)}. This observation contradicts thre existing plan the reactor
maintin and it is important here to relaize the temp[oral direction of
the causal link that is triggereing this inconsitency. Indeed at the
time we will observe this contradicting \texttt{At}, the rule that
generated this conlict is connected to the past observations
\texttt{Going} and \texttt{Go} which cannot be fully retracted without
potential global impact in the rest of the exisiting reactors --
implying a huge cost in backprogation that would be problematic.

More generally tying a token to future outcomes in a model that is meant
to be confronted to rela 
executionis something that is problematic. It may be valid ifg you can
gurantee that this will always happen (for example the reactor
managing \texttt{Agent.location} timeline guarantees that the outcome
of the \texttt{Going} will {\em always} result on ending at the traget
location) but should be manipulated while keeping in mind that
predicting the future is not absolute. In this case one could alter
this part of the model as follow in order to give the reactor more
flexibility:

@@

The alteration made to relax the constraint on linbe 5 from a strict 
\texttt{equals} to contains and then add the new constraint at line 11
that ensure that the last \texttt{Going} done during this \texttt{Go}
ends up in the location the agent wnats to be. This firts level of
relaxation will therefore allow the agent to usies multiple
\texttt{Going} should the first one fail to reach our objective.  The
probalem can even be relaxed further if needed but this relaxation
gicves already more felxibility for the reactor to recover from local
failures.

Synchronization remains here a deliberation process, and for this
reason there's always the risk that the search of a solution is not a
straightforward sequence of unit decision. As such process is required
by the architecture at every ticks for every reactors presently
running this need to be carefully designed. At the engine level we
made careful choices in term of both the focus of the search --
limited only to tokens that can potentially overlap the current tick
-- and its heuristic where the internal state flaw is considered last
during the search. The later proved to limit the amount of backtrack
in the search for a solution -- or at least how deep we need to
backtrack in our decision tree -- similarly the sequence in which we
evaluate the possible resolution for such a flaw -- presented
previously -- was selected according to simple assumptions that takes
benefit of a possibly already existing plan by assuming that if the
outcome of synchronization appears consistent with our current plan
then it probably reflects that this reactor plan is executed. Such
assumption helps direct the internal state evaluation and, in nominal
cases (where we assume that reactors are acting in concert), proved to
be a sound solution.

One final aspect to keep into account is that in \rx point of view
synchronization is the only critical task of a reactor. Indeed, such
process is used in order for the agent to identify its current state
and maintain it consistent throughout the agent life-time. A reactor
not being able to find a consistent solution during synchronization --
meaning that after evaluating all the possible solutions provided by
its model it did not found any plan that is not inconsistent -- not
only expresses that this reactor is not able anymore to explain the
state of the world but may jeopardize other reactors that relies on
its {\em internal} state. For this reason a failure to do
synchronization immediately result on the agent killing this reactor
and notifying any reactor that depends on its {\em internal} state
variables by posting the special observation \textsf{Failed}. By doing
so we allow for graceful degradation where other reactors can readapt
their plan as they receive this \textsf{Failed} observation.

>>>>

\item \emph{Partitioning} the global planning problem into multiple
  decision loops, called {\em reactors}. Each reactor has its own
  scope both functionally (one for example focusing on the conversion
  of high level waypoints into low-level commands) and temporally as
  each reactor declares its planning look-ahead and expected latency
  \comment{$\pi \&  \lambda$ both defined for the first time} before
  being able to produce plan to the agent depending on where it is
  located in a reactor dependency graph such as shown in
  Fig. \ref{fig:agent}.

@@

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=0.45\columnwidth]{figs/TREX_flow}
  \vskip-1cm
  \caption{\small Illustration of state information flow between the
    \texttt{Pilot} and the \texttt{Vehicle} reactors.  Observations
    transit in a bottom-up during synchronization and apply to the
    execution frontier $\tau$.  Goal information flow from the top to
    the bottom and concern the desired future values of state
    variables.\comment{ Fig should be cited in the text}}
  \label{fig:trex:flow}
\end{figure}

@@

\item \texttt{step} that accepts new goals requested (if any) on the
  {\em internal} timeline of a reactor and will execute one step of
  deliberation. In the context of \eu for example one deliberation
  step corresponds to a single flaw resolution (Section
  \ref{sec:europa:arch}) while ensuring that the resulting plan is not
  proven inconsistent. This implies that it also includes the
  backtracking of the search in such a case.  The returned value by
  the \texttt{step} function is a boolean which indicates whether the
  reactor needs an extra step to deliberate or if this last step
  resulted in a complete plan within the reactor's look-ahead. On the
  later, the part of the plan that concern this reactor {\em external}
  timelines is provided in order to be dispatched as goals to
  whichever reactor that is reponsible for these. \comment{By
    extending I may have saud things I developed below ... will have
    to check}
\end{itemize}

@@

We first show how a reactor can track and identify state when it does
not have any compelling need to deliberate with no future goals. Even
in this condition, \rx requires each reactor to produce at every tick
its {\em internal} state. For those reactors not owning {\em internal}
timelines, \rx makes sure that {\em external} observations are taken
into account for every tick as shown in \cite{py10}. Such a
requirement is enforced to ensure that all reactors share a consistent
view of the world with advancing time.

{\em External} state updates are managed by introducing these
observation in the plan structure as facts and ensure that they remain
compatible with the current plan mainainted by the reactor. The
challenging part of synchronization is to decide the {\em internal}
state values for the current tick. In the Shopping example for example
we have the following code snippet which helps identify when we can
\texttt{Buy} a product:

\begin{verbatim}
Agent::Buy {
  eq(duration, 10);
  // [...]
  contained_by(condition object.location.At currLocation);
  // [...]
}
\end{verbatim}

If \texttt{object.location} is {\em external} and the \texttt{Agent}
is {\em internal}, a robot may encounter the situation where it is at
the location where it can buy one (possibly among others)
product. This results om a decision problem related to the state value
of the \texttt{Agent} timeline where the robot can either {\em Buy} a
specific product now, wait a little longer before doing so or just
simply not buy any product at this location.

In the \eu based reactor this decision requirement is a new type of
flaw which forces full identification of internal state of the reactor
for the current tick. Using this new flaw we can describe the
synchronization process as:

@@

We introduce this requirement as a new type of flaw for the \eu
framework. This flaw forces the reactor to identify its internal state
for the current tick. By using this flaw one can describe the
synchronization process as:

@@

While this model is perfectly acceptable for planning it becomes
problematic when execution is taken into account especially as it puts
a strong emphasis between an evolving plan and an unknown future
(\texttt{Going} to a location and the {\em expected} future outcome of
ending at this location, ties it strongly to the constraint in line
$5$ above).  So for example, assume that the \texttt{Agent.location}
is {\em external} to the reactor which receives the observation
\texttt{Going(Home,SuperMarket)} which synchronization has resolved by
producing the observation \texttt{Go(Home, SuperMarket)} in the
\texttt{Agent} timeline. This would imply that the only possible
outcome on \texttt{Agent.location} is to be \texttt{At(SuperMarket)}
in the directly foreseeable future. In the real world the fact that we
are \texttt{Going} to a certain location does not necessarly imply
that we will succeed on ending up there. In one extreme case due to an
exogenous event, the road to the \texttt{SuperMarket} is closed which
could result in an alternate observation that the robot is
\texttt{At(7th st. \& Main)}. This observation contradicts the
existing plan the reactor is maintaining. Further, it is important to
understand the temporal direction of the \emph{causal link} that is
triggering the inconsistency. At the time the robot observes this
contradicting the \texttt{At} predicate, the rule that generated this
conflict is connected to past observations \texttt{Going} and
\texttt{Go} which cannot be easily retracted without a potential
global impact to the plan implying a huge cost in back-propagation
that would be problematic.

More generally tying a token to future outcomes in a model that is
meant to be confronted by execution is problematic. It might be valid
if one can guarantee that an outcome is predisposed to occur (for
example the reactor managing \texttt{Agent.location} timeline
guarantees that the outcome of the \texttt{Going} will {\em always}
result on ending at the traget location). However, given the very
nature of the planning problem and the ability to search outcomes and
replan, the model should be written so that a future projection can be
manipulated. %  while keeping in mind that predicting the future is not
% absolute. 
In this case one could alter this part of the model in order to give
the reactor more flexibility as follows:

@@

We do so by relaxing the constraint in line $5$ from a strict
\texttt{equals} to a \texttt{contains} and to add the new constraint
at line $11$ which ensures that the last \texttt{Going} ends up in the
location the agent actually wants to be at. This first level of
relaxation will therefore allow the agent to use multiple
\texttt{Going} tokens should the first one fail to reach the
objective.  % The problem can even be relaxed further if needed but
% this relaxation gives already more felxibility for the reactor to
% recover from local failures.
As the simple example above shows, synchronization involves a
deliberation process with the risk that the search for a solution is
not a straightforward sequence of unit decisions. Since such a process
is required by the architecture at \emph{every} tick for \emph{every}
reactor the model needs to be carefully designed.

At the engine level careful choices were made in both the focus of the
search -- limited only to tokens that can potentially overlap the
current tick -- and heuristics where the internal state flaw is
considered last during the search process. The later limits the amount
of backtracking for a solution -- or at least how deep one needs to
backtrack in a decision tree. Similarly the sequence in which we
evaluate the possible resolution for such a flaw -- presented
previously -- was selected according to some simple assumptions that % takes
% benefit of a possibly assume the existing plan by assuming that 
if the outcome of synchronization appears consistent with our current
plan then it probably reflects that the plan has executed
\comment{that the past is monotonic?}. Such an assumption helps direct
the internal state evaluation and in nominal cases, proves to be a
sound solution.

One final aspect to take into account is that from a \rx point of
view, synchronization is the only critical task of a reactor. Such a
process is used in order for the agent to identify its current state
and maintain it consistently throughout the agent's lifetime. For a
reactor not being able to find a consistent solution during
synchronization -- implying that after evaluating all possible
solutions provided by a model it was unable to find a consistent plan
-- not only indicates that a reactor is unable to explain the state of
the world, but more importantly it may jeopardize other reactors that
rely on its {\em internal} state. For this reason a failure to
synchronize immediately results in the agent killing a reactor and
notifying any reactor that depends on its {\em internal} state
variables; it does so by posting the special observation
\textsf{Failed}. And by doing so we allow for graceful degradation
where other reactors can readapt their plans as they receive this
observation.
